---
title: 'Homework 11: Advanced ggplotting'
author: "Natalia Aristizábal"
date: "12/11/2018"
output: html_document
---
Using the fake data set created in homework #7 to explore ggplot2. 
```{r}
library(ggplot2)
library(grid)
library("RColorBrewer")
  
# first re-create fake data from hw #7 & #8
forest <- rnorm(n=500,mean=10,sd=2) 
noise <- rnorm(n=500, sd=10) 
fruit <- 5.5 + forest*5 + noise

# add new variables to have more to play with in graphs
category <- rep(c("Ant", "No Ant"), each=250)  
site = rep(c("A","B","C","D","E"), each=100, times=5)   # added 5 sites
 
# put data into a dataframe
ID <- seq_len(500) 
fake_data <- data.frame(ID,forest,fruit,category, site)

# look at the data
head(fake_data)

myColors <- brewer.pal(6,"Set1")
names(myColors) <- levels(fake_data$category)
```


### Start making pretty graphs:

Exploring the data with a histogram
```{r}
p1 <- ggplot(data=fake_data, aes(x=fruit, fill= site)) +
  geom_histogram(binwidth = 10) +
  ylab("Count of fruit sizes") +
  xlab("Experimental site") + 
  ggtitle("Histogram") +
  theme_bw()

print(p1)
```

Showing all four variables (2 continuous and 2 categorical)
```{r}
# create a color palette
colors <- c("Ant" = "green", "No Ant" = "orange", "A" = "darkgreen", "B" = "orange", "C" = "brown", "D" = "pink", "E" = "blue")


p2 <- ggplot(fake_data, aes(forest, fruit), fill="category") +
  geom_point(aes(colour = factor(category), shape = factor(site))) +
   labs(
    x = "Forest cover in 1 km radius (%)",
    y = "Coffee fruit size (mg)",
    colour = "category",
    shape = "site") +
  scale_colour_manual(values = colors) +
  theme_classic() +
  ggtitle("Trends")
  
print(p2)
```

Another way of plotting all four variables 
```{r}
p3 <- ggplot(data=fake_data, aes(x=forest,y=fruit), col=site, fill=category) + 
  facet_grid(~site) +
  geom_boxplot(aes(y=fruit, lty=category), size=0.5) +
  ylab("Coffee fruit size (mg)") +
  xlab("Forest cover in 1 km radius (%)") + 
  theme_dark() +
  ggtitle("Dark Boxplot")
  
print(p3)
```


Now separating by categorical data on grids and colors
```{r}
p4 <- ggplot(data=fake_data, aes(x=forest,y=fruit, group=interaction(site), col=site )) + 
  facet_grid(~category) +
  geom_point(alpha = 0.3) + 
  ylab("Coffee fruit size (mg)") +
  xlab("Forest cover in 1 km radius (%)") + 
  theme_classic() +
  ggtitle("Classic Separation of Colors")
  
print(p4)
```

Trying different themes and colors
```{r}
# setting new color palette
myColors <- brewer.pal(9,"Set1")
names(myColors) <- levels(fake_data$category)

ggplot(data=fake_data, aes(x=forest,y=fruit), col=site, color=category) + 
  facet_grid(~site) +
  geom_violin(aes(y=fruit, color=category), size=0.5) +
  ylab("Coffee berry size (mg)") +
  xlab("Forest cover in 1 km radius (%)") + 
  scale_colour_manual(name="category", values=myColors) +
  theme_minimal() +
    ggtitle("Violin Distribution")

ggplot(data=fake_data, aes(x=forest,y=fruit), col=site, color=category) + 
  facet_grid(~site) +
  geom_jitter(aes(y=fruit, color=category), size=1) +
  scale_colour_manual(name="category", values=myColors) +
  ylab("Coffee berry size (mg)") +
  xlab("Forest cover in 1 km radius (%)") + 
  theme_gray() +
  ggtitle("My Own Categorical Colors and Gray Theme")

ggplot(data=fake_data, aes(x=forest,y=fruit), col=site, color=category) + 
  facet_grid(~site) +
  geom_jitter(aes(y=fruit, color=category), size=1) +
  scale_colour_manual(name="category", values=myColors) +
  ylab("Coffee berry size (mg)") +
  xlab("Forest cover in 1 km radius (%)") + 
  theme_void() +
  ggtitle("Void Theme")
```


Putting names to facet_wrap and moving the legend
```{r}
# Create labels
labs <- c("Largest farm","Second largest farm","Average size farm","Second Smallest farm","Smallest farm")
levels(fake_data$site) <- rev(labs)
head(fake_data)

p5 <- ggplot(fake_data, aes(x=forest,y=fruit), fill= category) +
  geom_jitter(aes(y=fruit, color=category), size=1) +
  stat_smooth(method=lm) +
  facet_wrap(~site) + 
  theme(
  legend.position = c(.95, .35),
  legend.justification = c("right", "top"),
  legend.box.just = "right",
  legend.margin = margin(6, 6, 6, 6)) +
  ylab("Coffee berry size (mg)") +
  xlab("Forest cover in 1 km radius (%)") + 
  ggtitle("Trends by Category and Site")

print(p5)
```


When making many plots, I could create an object with my own theme with publication standards to save space on coding later on. As such:
```{r}
mytheme <- theme(axis.title.x = element_text(size=12, color = "black", face = "bold"), 
                 axis.title.y = element_text(size=12, color = "black", face = "bold"),
                 axis.text = element_text(size=10, color ="black"),legend.title=element_blank(),
                 legend.position=c("right"),legend.background=element_rect(colour="black",size=0.4),
                 panel.background = element_rect(colour = "black", size=0.5, fill=NA)) +
  theme(legend.text=element_text(size=12)) 
```


Putting color into confidence intervals  
```{r}
p6 <- ggplot(data=fake_data, aes(x=forest,y=fruit), color=category) + 
  geom_jitter(aes(y = fruit), size = 0.5, alpha = 0.5) +
  stat_smooth(method=lm, se=TRUE, level = 0.95, colour= "pink", alpha= 0.8) +
  ylab("Coffee berry size (mg)") +
  xlab("Forest cover in 1 km radius (%)") + 
  geom_hline(yintercept=50, linetype="dashed") +
  mytheme +
  ggtitle("Clean Graph (and probably my favorite)")

print(p6)
```

We can put these graphs together side by side. 
```{r}
pushViewport(viewport(layout = grid.layout(3, 2)))
print(p1, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(p2, vp = viewport(layout.pos.row = 1, layout.pos.col = 2))
print(p3, vp = viewport(layout.pos.row = 2, layout.pos.col = 1))
print(p4, vp = viewport(layout.pos.row = 2, layout.pos.col = 2))
print(p5, vp = viewport(layout.pos.row = 3, layout.pos.col = 1))
print(p6, vp = viewport(layout.pos.row = 3, layout.pos.col = 2))  
```