---
title: 'Homework #6 Part II'
author: "Natalia Aristizábal"
date: "10/3/2018"
output: html_document
---

```{r, echo=TRUE}
#z <- read.table("MyDataFile.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
#str(z)#
#summary(z)

```


#### Open libraries

```{r, echo=TRUE}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

#### Read in data vector
To illustrate, we will generate some fake data here:

```{r, echo=TRUE}
# quick and dirty, a truncated normal distribution to work on the solution set
z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

#### Plot histogram of data

Plot a histogram of the data, using a modification of the code from lecture. Here we are switching from qplot to ggplot for more graphics options. We are also rescaling the y axis of the histogram from counts to density, so that the area under the histogram equals 1.0.

```{r, echo=TRUE}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

```{r, echo=TRUE}


```

```{r, echo=TRUE}


```

```{r, echo=TRUE}


```

```{r, echo=TRUE}


```