---
title: "Homework 8: Functions"
author: "Natalia Aristizábal"
date: "12/6/2018"
output: html_document
  
---
#### Using code from homework #7 to learn how to create functions in structured programming format.

Create a function to creake fake data, a function to analyse that data with a simple linear regression, and a function to graph the results of the analysis. 
```{r,echo=TRUE}
library(ggplot2)

#FUNCTION: create fake data
#INPUTS: parameters to generate fake data with normal distribution 
#OUTPUTS: data frame with data with 2 variables with normal distribution

create_fake_data <- function(size=100, xmean=10,xSD=2, noiseSD=10, slope=10) {
      forest <- rnorm(n=size,mean=xmean,sd=xSD) 
  noise <- rnorm(n=size, sd=noiseSD) 
      fruit <- 5.5 + forest*slope + noise
      # organize data into a dataframe
      ID <- seq_len(100) 
      fake_data <- data.frame(ID,forest,fruit)
      return(fake_data)
}
      
# check to see if function worked
fake_data <- create_fake_data()
head(fake_data)

#FUNCTION:basic regression analysis in R
#INPUTS:random data generated above 
#OUTPUTS: summary from linear model (lm)
regModel <- function(data=fake_data,x=fake_data$forest,y=fake_data$fruit) {
  reg <- lm(y ~ x)
  reg_sum <- summary(reg)
  return(reg_sum)
}

# check to see if function worked
regModel(create_fake_data())

#FUNCTION:basic graph to represent regression analysis 
#INPUTS:random data generated above 
#OUTPUTS: graphic representation of the linear model 
regGraph <- function() {
  p <- ggplot(data=fake_data, aes(x=forest,y=fruit)) +
                 geom_jitter() +
                 stat_smooth(method=lm)
  return(p)
}

# check to see if function worked
regGraph()
```

Now let's tweek one of the functions: 'regGraph'
```{r}
# changing the last function to graph to plot more information and an improved version of that graph with more details: 

#FUNCTION:advanced graph to represent regression analysis 
#INPUTS:random dataset generated with create_fake_data() 
#OUTPUTS: advanced graphic representation of the linear model 
regGraph <- function(){
  p <- ggplot(data=fake_data, aes(x=forest,y=fruit)) +
    geom_line(linetype="dotted", color="red", size=1) +
    geom_point(color="blue", size=0.5) +
    stat_smooth(method=lm) +
    labs(x = "Coffee fruit size (mg)",
         y = "Forest cover in 1 km radius (%)") +
  theme_classic() +
  ggtitle("Graph with more details") +
    theme(axis.title.x = element_text(size=20, color = "black", face = "bold"), 
                 axis.title.y = element_text(size=20, color = "black", face = "bold"),
                 axis.text = element_text(size=15, color ="black"),legend.title=element_blank(),
                 legend.position=c("right"),legend.background=element_rect(colour="black",size=0.4),
                 panel.background = element_rect(colour = "black", size=0.5, fill=NA)) +
  theme(legend.text=element_text(size=18)) 
  
  return(p)
}

# check to see if new function worked
regGraph()
```
Excellent! The functions are working and one function can feed into the next one. 
